---
title: "Bootstrapping methods"
author: "Xiaotai Chai"
date: "2/28/2018"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(boot)
```


1. A statistics student was interested in prices for used Mustang cars being offered for sale on an internet site. He sampled 25 cars from the website; the data are stored in the data set “MustangPrice” in the Lock5Data package. 

a.	Suppose we are interested in studying the correlation between Price and Miles for the cars. Describe how you might take one bootstrap sample from this data set. 
```{r}
#install.packages("Lock5Data")
library(Lock5Data)
attach(MustangPrice)

sample_corr <- function(data, i){
 sample = data[i,]
  correlation = cor(sample$Price,sample$Miles)
  return(correlation)
}

(result <- boot(MustangPrice, sample_corr, R=500 ))
# I might take one bootstrap sample by taking n amount of sample by replacement.
```

b.	What is the bootstrap estimate of the standard error?
```{r}
(std_error <- sd(result$t))
```

c.	Find a 95% bootstrap confidence interval for the correlation. Interpret the interval in context.
```{r}
(CI_corr <- boot.ci(result))
## Interpretation: We are confident that 95% of the correlation would fall between -0.9 and -0.7. There's a negative correlation between price and miles. When miles go up, price goes down.
```

d.	Suppose we would like to create a 95% bootstrap confidence interval for the median price of the mustangs. Simulate 1000 bootstrap samples and graph the bootstrap distribution for the sample median price. If appropriate, construct a 95% bootstrap CI for the median price; if not appropriate, explain why.
```{r}
sample_med <- function(data, i){
 sample = data[i,]
  med = median(sample$Price)
  return(med)
}

(result2 <- boot(MustangPrice, sample_med, R=1000 ))
plot(result2)
(CI_med <- boot.ci(result2))
## It's appropriate.
```




