---
title: "Relationship"
author: "Xiaotai Chai"
date: "11/6/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(ggplot2)
library(tidyverse)
```


# Set working directory  
```{r}
# double check that this puts you in the right dir
project.dir <- getwd()
setwd(project.dir)
dataset.dir <- "Data"
outputs.dir <- "Output"
```

# Read in cleaned csv file
```{r}
# todo replace with fread
outcomes2013 <- fread("filtered_2013.csv", na.strings=c("<NA>","PrivacySuppressed"))
#head(outcomes2013)
```

# Find relationship between completion within 4 years at original institution and Gender of student
```{r}
female_complete4yr <- as.numeric(outcomes2013$FEMALE_COMP_ORIG_YR4_RT)
male_complete4yr <- as.numeric(outcomes2013$MALE_COMP_ORIG_YR4_RT)
num_completed_by_gender <- na.omit(data.frame(female_complete4yr,male_complete4yr))  # omit rows were either F or M is NA

plot(num_completed_by_gender, xlab="# females graduated", ylab="# males graduated")+title("# of graduates, male vs female")

lm.fit <- lm(female_complete4yr ~ male_complete4yr, data = num_completed_by_gender)
summary(lm.fit)
plot(lm.fit)

MEAN <- c(mean(as.numeric(outcomes2013$MALE_COMP_ORIG_YR4_RT),na.rm=TRUE), mean(as.numeric(outcomes2013$FEMALE_COMP_ORIG_YR4_RT),na.rm=TRUE)) 

STD <- c(sd(as.numeric(outcomes2013$MALE_COMP_ORIG_YR4_RT),na.rm=TRUE), sd(as.numeric(outcomes2013$FEMALE_COMP_ORIG_YR4_RT),na.rm=TRUE)) 

MEDIAN <- c(median(as.numeric(outcomes2013$MALE_COMP_ORIG_YR4_RT),na.rm=TRUE), median(as.numeric(outcomes2013$FEMALE_COMP_ORIG_YR4_RT),na.rm=TRUE)) 

mydata <- data.frame(MEAN, STD, MEDIAN)
row.names(mydata) <- c("male_complete4yr", "female_complete4yr")
mydata
```
```

# Comparing NCF to other colleges
```{r}
ncf <- outcomes2013[outcomes2013$INSTNM == "New College of Florida",]

female_percent <- outcomes2013[,c("INSTNM","FEMALE")]
colnames(female_percent) <- c("School", "Value")
female_percent <- as.data.frame(mutate(female_percent, Feature="Percnt_female"))
female_percent$Value <- as.numeric(female_percent$Value)
female_percent$School[which(female_percent$School != "New College of Florida")] = "Other schools"
ggplot(female_percent, aes(y=Value, x=Feature, group=School, fill=School)) + geom_boxplot(position=position_dodge(1))
```

New College is on the lower end of percent female.

```{r}
percent_recv_loans <- outcomes2013[,c("INSTNM","PCTFLOAN")]
colnames(percent_recv_loans) <- c("School", "Value")
percent_recv_loans <- as.data.frame(mutate(percent_recv_loans, Feature="Percnt_reciv_loan"))
percent_recv_loans$Value <- as.numeric(percent_recv_loans$Value)
percent_recv_loans$School[which(percent_recv_loans$School != "New College of Florida")] = "Other schools"
ggplot(percent_recv_loans, aes(y=Value, x=Feature, group=School, fill=School)) + geom_boxplot(position=position_dodge(1))
```

New College is in the lower tail of recieving loans.

```{r}
percnt_older_than_25yr <- outcomes2013[,c("INSTNM","UG25ABV")]
colnames(percnt_older_than_25yr) <- c("School", "Value")
percnt_older_than_25yr <- as.data.frame(mutate(percnt_older_than_25yr, Feature="Percnt_older_than_25yr"))
percnt_older_than_25yr$Value <- as.numeric(percnt_older_than_25yr$Value)
percnt_older_than_25yr$School[which(percnt_older_than_25yr$School != "New College of Florida")] = "Other schools"
ggplot(percnt_older_than_25yr, aes(y=Value, x=Feature, group=School, fill=School)) + geom_boxplot(position=position_dodge(1))
```

New College had an insignificant number of >25 year olds.

```{r}
# avg family income
Avg_fam_income <- outcomes2013[, c("INSTNM","FAMINC")]
colnames(Avg_fam_income) <- c("School", "Value")
Avg_fam_income$Value <- as.numeric(Avg_fam_income$Value)
Avg_fam_income <- as.data.frame(mutate(Avg_fam_income, Feature="Avg_fam_income"))
Avg_fam_income$School[which(Avg_fam_income$School != "New College of Florida")] = "Other schools"
ggplot(Avg_fam_income, aes(y=Value, x=Feature, group=School, fill=School)) + geom_boxplot(position=position_dodge(1))
```

New College kids come from rich families.


```{r}
mean_income_graduates_after_8yr <- outcomes2013[, c("INSTNM","MN_EARN_WNE_P8")]
colnames(mean_income_graduates_after_8yr) <- c("School", "Value")
mean_income_graduates_after_8yr <- as.data.frame(mutate(mean_income_graduates_after_8yr, Feature="Mean_income_graduates_after_8yr"))
mean_income_graduates_after_8yr$Value <- as.numeric(mean_income_graduates_after_8yr$Value)
mean_income_graduates_after_8yr$School[which(mean_income_graduates_after_8yr$School != "New College of Florida")] = "Other schools"
ggplot(mean_income_graduates_after_8yr, aes(y=Value, x=Feature, group=School, fill=School)) + geom_boxplot(position=position_dodge(1))
```

New College is around the mean income of graduates.

```{r}
# Share of students earning over $25,000/year (threshold earnings) 8 years after entry
perc_graduates_earning_grtrthn25k_peryer <- outcomes2013[,c("INSTNM","GT_25K_P8")]
colnames(perc_graduates_earning_grtrthn25k_peryer) <- c("School", "Value")
perc_graduates_earning_grtrthn25k_peryer <-  as.data.frame(mutate(perc_graduates_earning_grtrthn25k_peryer, Feature="Perc_graduates_earning_grtrthn25k_peryer"))
perc_graduates_earning_grtrthn25k_peryer$School[which(perc_graduates_earning_grtrthn25k_peryer$School != "New College of Florida")] = "Other schools"
perc_graduates_earning_grtrthn25k_peryer$Value <- as.numeric(perc_graduates_earning_grtrthn25k_peryer$Value)
ggplot(perc_graduates_earning_grtrthn25k_peryer, aes(y=Value, x=Feature, group=School, fill=School)) + geom_boxplot(position=position_dodge(1))
```

NCF graduates make a little less than averge than other graduates.

```{r}
# instructional expenditures
institutional_expenditures <- outcomes2013[,c("INSTNM","INEXPFTE")]
colnames(institutional_expenditures) <- c("School", "Value")
institutional_expenditures <-  as.data.frame(mutate(institutional_expenditures, Feature="institutional_expenditures"))
institutional_expenditures$School[which(institutional_expenditures$School != "New College of Florida")] = "Other schools"
institutional_expenditures$Value <- as.numeric(institutional_expenditures$Value)
ggplot(institutional_expenditures, aes(y=Value, x=Feature, group=School, fill=School)) + geom_boxplot(position=position_dodge(1))
```

New College is in the lower tail of institutional expenditure spendings


```{r}
# Net tuition revenue per full-time equivalent student
tuition_in_state <- outcomes2013[,c("INSTNM","TUITFTE")]
colnames(tuition_in_state) <- c("School", "Value")
tuition_in_state <-  as.data.frame(mutate(tuition_in_state, Feature="tuition_in_state"))
tuition_in_state$Value <- as.numeric(tuition_in_state$Value)
tuition_in_state$School[which(tuition_in_state$School != "New College of Florida")] = "Other schools"
ggplot(tuition_in_state, aes(y=Value, x=Feature, group=School, fill=School)) + geom_boxplot(position=position_dodge(1))
```
About average.


```{r}
# The median original amount of the loan principal upon entering repayment
loan_principal <- outcomes2013[,c("INSTNM","DEBT_MDN")]
colnames(loan_principal) <- c("School", "Value")
loan_principal <-  as.data.frame(mutate(loan_principal, Feature="loan_principal"))
loan_principal$Value <- as.numeric(loan_principal$Value)
loan_principal$School[which(loan_principal$School != "New College of Florida")] = "Other schools"
ggplot(loan_principal, aes(y=Value, x=Feature, group=School, fill=School)) + geom_boxplot(position=position_dodge(1))
```

Way lower loan principal!



```{r}
# 3 yr cohort default rate
default_rate <- outcomes2013[,c("INSTNM","CDR3")]
colnames(default_rate) <- c("School", "Value")
default_rate <-  as.data.frame(mutate(default_rate, Feature="default_rate"))
default_rate$Value <- as.numeric(default_rate$Value)
default_rate$School[which(default_rate$School != "New College of Florida")] = "Other schools"
ggplot(default_rate, aes(y=Value, x=Feature, group=School, fill=School)) + geom_boxplot(position=position_dodge(1))
```

I thought we'd have higher because we're communist hippies, but we are in the lower tail

